---
title: 'Fuel analysis for Oak Symposium: Downed woody debris'
author: "Morgan Gray"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries -----
rm(list = ls())
library(tidyverse)   ## To manipulate data frames
library(here)   ## To manage directories
library(janitor)   ## To clean data frames
library(rstatix)  ## For repeated measure ANOVA
library(broom)  ## To format output tables from statistical tests
library(lemon)  ## To manipulate faceted ggplots
library(kableExtra)  ## For tables in rmarkdown
library(ggpubr)  ## For qq plots

# Define file paths ----
path_in <- here("input")
path_cache <- here(path_in, "cache")
path_lookup <- here(path_in, "lookup-tables")
path_out <- here("analysis/tubbs")
path_out_plot <- here(path_out, "plots")

# Define helpers ----
date_today <- Sys.Date()

# lookup_dwd <- read_csv(here(path_in, "lookup_variables.csv")) %>%
#   filter(data_type %in% "dw") %>%
#   select(fuel_class = name, 
#          lab_fuel = label)

# Define plot colors ----
plot_colors <- 
  read_csv(here(path_in, 
                "plot-colors.csv")) %>%
  arrange(palette, palette_subset, levels)

colors_fuel <- 
  plot_colors %>%
  filter(palette %in% "fuel_class", 
         palette_subset %in% "all")

# colors_dw <- c("#7da8b0", "#c4c4c4")
# 
# colors_dw5 <- 
#   plot_colors %>%
#   filter(palette_subset %in% "dw") %>%
#   pull(hex_code)

# colors_year <- 
#   plot_colors %>%
#   filter(palette %in% "year", 
#          palette_subset %in% "1pre_3post") %>%
#   pull(hex_code) 


# ========================================================== -----
# WRITE FUNCTIONS ----
# Basic functions ----
#   %nin% ----
"%nin%" <- Negate("%in%")
#   fxn_digit ----
fxn_digit <- function(x){
  as.numeric(format(round(x, 3), nsmall = 3))
}
#   fxn_signif ----
fxn_signif <- function(df){
  df %>%
    mutate(p_adj_sig = case_when(
      p_adj < 0.001 ~ "***", 
      p_adj > 0.001 & p_adj < 0.01 ~ "**", 
      p_adj > 0.01 & p_adj < 0.05 ~ "*", 
      p_adj > 0.05 ~ "n.s.")) %>%
    arrange(p_adj)
    
}

# Data transformation (normalize and standardize) ----
#   fxn_tranform_ordnorm ----
fxn_tranform_ordnorm <- function(){
  
  datalist_normalized <- list()
  for(index_class in list_classes){
    
    subset <- 
      input_class %>%
      filter(fuel_class %in% index_class) %>%
      rename(value_raw = value) 
    
    bn_o <- bestNormalize::orderNorm(subset$value_raw, standardize = TRUE)
   
    datalist_normalized[[index_class]] <- 
        subset %>%
        mutate(value_ordnorm = bn_o$x.t) %>%
        select(fuel_class, year, plot_id, starts_with("value"), starts_with("lab")) 
   
    }
     
  bind_norm <- do.call(bind_rows, datalist_normalized)

}

#   fxn_transform_arcsine ----
fxn_transform_arcsine <- function(index_class){
  
  subset <- 
    input_class %>%
    filter(fuel_class %in% index_class)  %>%
    rename(value_raw = value) 
  
  bn <- bestNormalize::arcsinh_x(subset$value_raw, standardize = TRUE)
  
  normalized <-
    subset %>%
    mutate(value = bn$x.t, 
           transform = "arcsine") %>%
    relocate(value, .after = value_raw) %>%
    relocate(transform, .after = value)
}
#   fxn_transform_log ----
fxn_transform_log <- function(index_class){
  
  subset <- 
    input_class %>%
    filter(fuel_class %in% index_class)  %>%
    rename(value_raw = value) 
  
  bn <- bestNormalize::log_x(subset$value_raw, standardize = TRUE)
  
  normalized <-
    subset %>%
    mutate(value = bn$x.t, 
           transform = "log") %>%
    relocate(value, .after = value_raw) %>%
    relocate(transform, .after = value)
  
}
# Statistical tests ----
#   fxn_aov_me ----
fxn_aov_me <- function(df){
  df %>%
    anova_test(dv = value, 
               wid = plot_id, 
               within = year) %>%
    get_anova_table() %>%
    adjust_pvalue(method = "bonferroni") %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(method = "me", 
           data_type = index_type) %>%
    rename(statistic = f) %>% 
    relocate(data_type, 
           method, 
           effect, 
           starts_with("p_adj"), 
           statistic,
           starts_with("d"), 
           ges)
}

#   fxn_pwc ----
# Excludes fuel_class, is for pwc of fuel_type only
fxn_pwc <- function(df){
  df %>%
    pairwise_t_test(
      value ~ year, 
      paired = TRUE,
      p.adjust.method = "bonferroni") %>%
    clean_names() %>%
    mutate(method = "pwc", 
           data_type = index_type, 
            statistic = fxn_digit(statistic)) %>%
    rename(p_adj_sig = p_adj_signif) %>%
    select(data_type, 
             method, 
             starts_with("group"),
             starts_with("p_adj"), 
             statistic,
             starts_with("d"),  
             starts_with("p"))
}

#   fxn_aov2_me ----
fxn_aov2_me <- function(df){
  df %>%
    anova_test(dv = value, 
               wid = plot_id, 
               within = c(fuel_class, year)) %>%
    get_anova_table() %>%
    adjust_pvalue(method = "bonferroni") %>%
    as_tibble() %>%
    clean_names() %>%
    mutate(method = "aov2", 
           data_type = index_type)  %>%
    rename(statistic = f) %>%
    select(data_type, 
             method, 
             effect, 
             starts_with("p_adj"), 
             statistic,
             starts_with("d"), 
             ges)
}

#   fxn_pwc2 ----
# Includes a column for fuel_class
fxn_pwc2 <- function(df){
  df %>%
    pairwise_t_test(
      value ~ year, 
      paired = TRUE,
      p.adjust.method = "bonferroni") %>%
    clean_names() %>%
    mutate(method = "pwc", 
           data_type = index_type, 
           statistic = fxn_digit(statistic)) %>%
    rename(p_adj_sig = p_adj_signif) %>%
    select(data_type, 
             method, 
             fuel_class,
             starts_with("group"),
             starts_with("p_adj"), 
             statistic,
             starts_with("d"))

 
}
# Summary plots and tables ----
#   fxn_kable ----
fxn_kable <- function(df){
  df  %>%
    knitr::kable() %>%
    kable_styling(bootstrap_options = c("striped", "hover"), 
                  full_width = F,  
                  position = "left", 
                  fixed_thead = T)
}
  
#   fxn_posthoc_plot ----
fxn_posthoc_plot <- function(index_subset, index_input){
  
# Create annotations for boxplot ----
annotation <- 
  index_input %>%
  filter(fuel_class %in% index_subset) %>%
  pairwise_t_test(value ~ year, paired = TRUE, p.adjust.method = "bonferroni")  %>%
  add_xy_position(x = "year") 

title <- 
  index_input %>%
  filter(fuel_class %in% index_subset) %>%
  distinct(lab_fuel) %>%
  pull()
  
# Create boxplot comparing years  ----
index_input %>%
  filter(fuel_class %in% index_subset) %>%
   ggboxplot(x = "year", 
            y = "value", 
            fill = "lab_year_abbr", 
            palette = colors_year, 
            outlier.size = 0.3,
            width = 0.6) +
  ylab("Mean tons per acre") + 
  labs(title = title, 
       fill = "Year") +
   theme(axis.title.x = element_blank(), 
        legend.position = "top")  + 
  # Add p-values
  stat_pvalue_manual(annotation, tip.length = 0, hide.ns = TRUE) 

}
#   fxn_posthoc_plot_bt ----
fxn_posthoc_plot_bt <- function(index_subset, index_input){
  
# Create annotations for boxplot ----
annotation <- 
  index_input %>%
  filter(fuel_class %in% index_subset) %>%
  pairwise_t_test(value ~ year, paired = TRUE, p.adjust.method = "bonferroni")  %>%
  add_xy_position(x = "year") 

title <- 
  index_input %>%
  filter(fuel_class %in% index_subset) %>%
  distinct(lab_fuel) %>%
  pull()
  
# Create boxplot comparing years  ----
index_input %>%
  filter(fuel_class %in% index_subset) %>%
  ggboxplot(x = "year", 
            y = "value_raw", 
            fill = "lab_year_abbr", 
            palette = colors_year, 
            outlier.size = 0.3,
            width = 0.6) +

  theme(legend.position = "top", 
        # axis.title.x = element_blank(), 
        legend.text=element_text(size=8),
        axis.text = element_text(size = 8))  + 
  
  labs(title = title, 
       fill = "Year", 
       x = "Year",
       y = "Mean tons per acre") +
  
  # Add p-values
  stat_pvalue_manual(annotation, tip.length = 0, hide.ns = TRUE) 

}
#   fxn_posthoc_table ----
fxn_posthoc_table <- function(index_subset, index_input){
  
title <- 
  index_input %>%
  filter(fuel_class %in% index_subset) %>%
  distinct(lab_fuel) %>%
  pull()
  
index_input %>%
 filter(fuel_class %in% index_subset) %>%
  fxn_pwc() %>%
    mutate(fuel_class = title) %>%
    relocate(data_type, 
             fuel_class, 
             starts_with("group"), 
             starts_with("p_adj"), 
             statistic, 
             starts_with("d")) %>%
  select(-p) %>%
  arrange(p_adj) %>%
  fxn_kable()
}

# ========================================================== -----
# Read input data ----
input_data <- 
  read_csv(here(path_in, "fuel_for-aov_2022.csv")) %>%
  mutate(value = fxn_digit(value)) %>%
  relocate(c(statistic, subset), .after = value)  

dw_class <-
  input_data %>%
  filter(data_type %in% "dw", 
         statistic %in% "mean")

dw_total <-
  input_data %>%
  filter(data_type %in% "dw", 
         statistic %in% "total") %>%
  remove_empty("cols")

list_classes_dw <- unique(dw_class$fuel_class)

gf_class <-
  input_data %>%
  filter(data_type %in% "gf", 
         statistic %in% "mean")

gf_total <-
  input_data %>%
  filter(data_type %in% "gf", 
         statistic %in% "total") %>%
  remove_empty("cols")

list_classes_gf <- unique(gf_class$fuel_class)


```

------------------------------------------------------------------------

## Overview

The aim of this research was to understand the impact of fuel load management (i.e., thinning) on downed woody debris, duff, and litter. Our objective was to evaluate the short-term impact of thinning on levels of downed woody debris (1-hr, 10-hr, 100-hr, 1000-hr rotten, 1000-hr sound) and duff and litter. To meet this objective, we used a "treatment and control" approach to compare fuel levels before (control) and after (treatment) thinning was applied at five vegetation monitoring plots.

The control and treatment data sets were collected at two time points in 2021. Control data were collected before thinning, between 6/9/21 and 7/7/21. Treatment data were collected after thinning on 10/18/21 and 10/19/21, with the exception of measurements for plot FOR10 which surveyed on 12/9/21. On average, 133.6 days elapsed between control and treatment surveys (range: 104 to 183, standard deviation = 27.2).

The data for each time point were collected by conducting surveys along three transects within each of the five plots. Each transect had one quadrat for downed woody debris (n = 3 quadrats per plot) and two quadrats for ground fuels (n = 6 quadrats per plot). Each time point had 15 data points for each downed woody debris fuel class (5 plots x 3 transects x 1 quadrat) and 30 data points each for duff and litter (5 plots x 3 transects x 2 quadrats).

We used these data to answer the following questions:

1.  Did total downed wood debris differ between control and treatment surveys?

2.  Did fuel load for any of the five fuel classes differ between control and treatment surveys? 

3.  Did the total fuel load (all duff and litter combined) differ between control and treatment surveys?

4.  Did fuel load for duff and/or litter differ between control and treatment surveys?  

#### Statistical methods

Statistically significant differences were identified using repeated measures analysis of variance (ANOVA) and post hoc comparisons. P-values were adjusted using the Bonferroni multiple testing correction method.

[NOTE: A repeated measures ANOVA is appropriate to use for a paired time series (before vs. after, treatment vs. control). For these data in particular, the approach accounts for elements that structure the data set in addition to time. For example, transects within each plot are more likely to be similar than transects between plots.]

The repeated-measures ANOVA is used for analyzing data where same subjects are measured more than once. This test is also referred to as a within-subjects ANOVA or ANOVA with repeated measures. The "within-subjects" term means that the same individuals (here, individuals are plots) are measured on the same outcome variable under different time points. The main goal of a repeated measures ANOVA is to evaluate if there is a statistically significant interaction effect between within-subjects factors in explaining a continuous outcome variable. The repeated measures ANOVA makes the following assumptions about the data:

-   No significant outliers in any cell of the design
-   Normality: the outcome (or dependent) variable should be approximately normally distributed in each cell of the design
-   Assumption of sphericity: the variance of the differences between groups should be equal

We assessed outliers using the the interquartile range (IQR; IQR = Q3 - Q1). Values above Q3 + 1.5xIQR or below Q1 - 1.5xIQR are considered as outliers. Values above Q3 + 3xIQR or below Q1 - 3xIQR are considered as extreme points (or extreme outliers). Q1 and Q3 are the first and third quartile, respectively. Extreme outliers can be due to data entry errors, measurement errors, or unusual values. The outlier may be included if one believes the result will not be substantially affected; this can be evaluated by comparing the result of the ANOVA with and without the outlier.

We assessed normality by visual inspection of a QQ plot for each time point. A QQ plot draws the correlation between a given data and the normal distribution. We also conducted the Shapiro-Wilk test for each time point. Using this method, normally distributed data will have p-value \>0.05.

The assumption of sphericity was checked during the computation of the ANOVA test using the R function anova_test() [rstatix package]. The Mauchly's test was internally used to assess the sphericity assumption, and the Greenhouse-Geisser sphericity correction was automatically applied to factors violating the sphericity assumption.

#### Data frames

For each fuel type (downed woody debris, ground fuels), we compared the plot-level values between control and treatment surveys. We defined "total (or all) downed woody debris" as the combined total of the five fuel classes within a plot: 1-hr, 10-hr, 100-hr, 1000-hr rotten, 1000-hr sound. We defined "total (or all) ground fuels" as the combined total of the duff and litter fuel classes within a plot.

Recall that surveys were conducted along three transects within each plot; each transect had one quadrat for downed woody debris (n = 3 quadrats per plot) and two quadrats for ground fuels (n = 6 quadrats per plot). To account for replicate surveys within each plot, we used the plot-level mean for each fuel class as input for **... [TO DO WHAT?].** Below is a subset of the data frame showing the plot-level means by time point.

```{r, echo=FALSE, message=FALSE}
gf_class %>%
  arrange(plot_id, fuel_class, desc(timing)) %>%
  select(fuel_type = lab_type, 
         fuel_class = lab_fuel, 
         plot_id, 
         timing,
         value) %>%
   spread(timing, value) %>%
  relocate(pre, .before = post) %>%
  head() %>%
  fxn_kable()  

```

To determine the plot-level total by fuel type at each time point, we summed the plot means for all fuel classes by fuel type Below is a subset of the data frame showing the plot-level totals by time point.

```{r, echo=FALSE, message=FALSE}
dw_total %>%
  arrange(plot_id, desc(timing)) %>%
  select(fuel_type = lab_type, 
         plot_id, 
         timing,
         value) %>%
  spread(timing, value) %>%
  relocate(pre, .before = post) %>%
    head() %>%
  fxn_kable()

```

------------------------------------------------------------------------

## Q1: Did total downed woody debris differ between control and treatment surveys? 

The first question we asked was whether there was a significant main effect of thinning treatment on total fuel load (plot-level mean) for each fuel type. A one-way repeated measures ANOVA was used to determine whether the mean fuel load was significantly different between the control and treatment surveys.

### Summary statistics for the plot-level total amount of downed woody debris by year

The following table summarizes the plot-level total amount of downed woody debris for each time point.

```{r, echo=FALSE}
input_total %>%
  group_by(year) %>%
  get_summary_stats(value, type = "mean_sd")  %>%
  mutate(data_type = index_type,
         fuel_class = "All") %>%
  select(data_type, fuel_class, year, mean, sd, n) %>%
     fxn_kable()

```

### Visualization of the plot-level total amount of downed woody debris by year

A boxplot of total downed woody debris by year showed lower values in post-fire years. Specifically, we observed a decrease in the depth of all downed woody debris immediately following the fire (2017; 0-y post-fire). Variance in 2017 was much greater than at any other time point. The mean depth of downed woody debris did not return to pre-fire levels in subsequent years. However, the fuel load increased from 2017 levels by 2019, and showed little change between 2019 and 2021.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p_total <- 
  input_total %>%
  ggboxplot(x = "lab_year", 
            y = "value", 
            fill = "year", 
            outlier.size = 0.3,
            palette = colors_year) + 
  ylab("Mean tons per acre") + 
  theme(legend.position = "none", 
        axis.title.x = element_blank()) + 
  scale_x_discrete(labels = function(lab_year) str_wrap(lab_year, width = 20)) + 
  labs(title = paste0(index_type, " (all)")) 

p_total


```

### Check assumptions

##### *Check for outliers*

There were no extreme outliers in the data set.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
input_total %>%
  group_by(year) %>%
  identify_outliers(value)  %>%
  mutate(data_type = index_type,
         fuel_class = "All") %>%
  clean_names() %>%
  select(data_type, fuel_class, year, plot_id, value, 
         starts_with("is")) %>%
    # filter(is_extreme == TRUE) %>%
     fxn_kable()

```

##### *Check for normality*

Plot-level values for downed woody debris were not normally distributed for three of the four time points (2016, 2017, 2019), as assessed by Shapiro-Wilk's test.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

input_total %>%
  select(id = plot_id, 
         time = year, 
         score = value) %>%
  group_by(time) %>%
  shapiro_test(score)  %>%
    clean_names() %>%
  mutate(data_type = index_type,
         fuel_class = "All", 
         statistic = fxn_digit(statistic), 
         is_normal = p>0.05) %>%
  select(data_type, 
         fuel_class, 
         year = time, 
         is_normal,
         p, 
         statistic) %>%
  arrange(is_normal, year) %>%
      filter(is_normal == FALSE) %>%
     fxn_kable()

```

Several points in the QQ plot fell outside the reference range.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggqqplot(input_total, "value", 
         facet.by = "year", 
         color = "year",
         palette = colors_year)

```

##### *Normalize data and repeat checks*

We applied an arcsine transformation to normalize the plot-level values for total downed woody debris, calculated as log(x + sqrt(x\^2 + 1)). The arcsine transformation (also called the arcsine square root transformation, or the angular transformation) was identified as the most suitable method using the R function bestNormalize() [bestNormalize package]. Values were standardized upon normalization to have a mean of 0 and standard deviation of 1.

The transformed values for total downed woody debris were normally distributed at each time point (p \> 0.05), as assessed by Shapiro-Wilk's test.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

bn_total <- bestNormalize::arcsinh_x(input_total$value, standardize = TRUE)

input_total_transform <- 
  input_total %>%
  mutate(value = bn_total$x.t) 

input_total_transform %>%
  select(id = plot_id, 
         time = year, 
         score = value) %>%
  group_by(time) %>%
  shapiro_test(score)  %>%
    clean_names() %>%
  mutate(data_type = index_type,
         fuel_class = "All", 
         statistic = fxn_digit(statistic), 
         is_normal = p>0.05) %>%
  select(data_type, 
         fuel_class, 
         year = time, 
         is_normal,
         p, 
         statistic) %>%
  arrange(is_normal, year) %>%
  # filter(is_normal == FALSE) %>%
     fxn_kable()

```

All the points on the below QQ plot fell approximately along the reference line, we could assume normality.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggqqplot(input_total_transform, "value", 
         facet.by = "year", 
         color = "year",
         palette = colors_year)
```

The two outliers (not extreme) detected in the untransformed data were absent after the transformation was applied.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

input_total_transform %>%
  group_by(year) %>%
  identify_outliers(value)  %>%
  mutate(data_type = index_type,
         fuel_class = "All") %>%
  clean_names() %>%
  select(data_type, fuel_class, year, plot_id, value, 
         starts_with("is")) %>%
    # filter(is_extreme == TRUE) %>%
     fxn_kable()

```

### Repeated measures ANOVA test

We found no significant main effect of year on the plot-level total for downed woody debris (p = 0.359).

```{r, echo=FALSE, warning=FALSE}
  # Effect of year (all fuel classes)
input_total_transform %>%
    fxn_aov_me() %>%
  fxn_signif() %>%
    mutate(fuel_class = "All") %>%
  clean_names() %>%
  select(-p, 
         -p_05) %>%
   relocate(data_type, 
            fuel_class,
             effect, 
             starts_with("p_adj"), 
             statistic, 
             starts_with("d_"), 
             ges) %>%
    fxn_kable()


```

### Post hoc tests

We conducted post hoc pairwise comparisons between the levels of the within-subjects factor (here, year). The result of paired t-tests between years showed no significant difference in fuel load between years at a significance level of \< 0.05. This finding is consistent with the lack of main effect found for year.

The results from the pairwise comparisons are shown below as (1) a boxplot of total downed woody debris by year (the lack of p-values reflects the absence of significant comparisons), and (2) a table of test results.

```{r, echo=FALSE}
pwc <- input_total_transform %>%
  pairwise_t_test(value ~ year, paired = TRUE, p.adjust.method = "bonferroni")  %>% 
  add_xy_position(x = "year")

aov_me <- input_total_transform %>%
    anova_test(dv = value, wid = plot_id, within = year)
  
p_total + 
  labs(subtitle = get_test_label(aov_me, detailed = TRUE),
       caption = get_pwc_label(pwc)) +
  stat_pvalue_manual(pwc, hide.ns = TRUE) 

```

```{r, echo=FALSE, warning=FALSE}
  # Effect of year (all fuel classes)
input_total_transform %>%
    fxn_pwc() %>%
    mutate(statistic = fxn_digit(statistic),
           data_type = index_type,
         fuel_class = "All") %>%
  select(-p) %>%
  relocate(data_type, 
           fuel_class,
             starts_with("group"), 
             starts_with("p_adj"), 
             statistic, 
             starts_with("d")) %>%
  arrange(p_adj) %>%
    fxn_kable()

```

------------------------------------------------------------------------

## Q2: Did downed woody debris for control and treatment surveys differ by fuel class?

Next, we investigated whether there was a significant change in plot-level fuel load over time when accounting for fuel class. A two-way repeated measures ANOVA was used to determine whether there was a significant interaction between year and fuel class on fuel load.

Here, the effect of year on fuel load was our focal variable of primary concern. However, the effect year may differ between fuel classes, so fuel_class was considered a moderator variable.

### Summary statistics for downed woody debris by year and fuel class

The following table summarizes downed woody debris by class for each time point.

```{r, echo=FALSE}
input_class %>%
  group_by(year, fuel_class) %>%
  get_summary_stats(value, type = "mean_sd")  %>%
  arrange(fuel_class, year) %>%
  mutate(data_type = index_type) %>%
  select(data_type, fuel_class, year, mean, sd) %>%
  gather(statistic, value, mean:sd) %>%
  spread(year, value) %>%
     fxn_kable()

```

### Visualization of downed woody debris by year and fuel class

A boxplot of downed woody debris by year and fuel class showed the post-fire trends differed among the five downed woody debris fuel classes. [Note: The y-axis scale in the figure below differs by fuel class]

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p_class <-
  input_class %>%
  bind_rows(input_total) %>%
  # Configure factors so total ("all") appears first in facet
  mutate(fuel_class = ifelse(is.na(fuel_class), "all dw", fuel_class), 
         lab_fuel = as.character(str_replace(lab_fuel, "hr", "h")), 
         yr_nn = str_sub(year, 3, 4)) %>%
  arrange(plot_id, year, fuel_class) %>%
  mutate(fuel_class = as_factor(fuel_class), 
         lab_fuel = as_factor(lab_fuel)) %>%
  ggboxplot(x = "yr_nn", 
            y = "value", 
            fill = "lab_year_abbr",
            outlier.size = 0.3,
            palette = colors_year) + 
  theme(legend.position = "top", 
        # axis.title.x = element_blank(), 
        # axis.text.x = element_blank(),
        axis.text = element_text(size = 8),
        panel.spacing = unit(0.5, "lines"),
        # Reduce font size for facet strip text
        legend.text=element_text(size=8),
        strip.text = element_text(size = 8),
        strip.background = element_rect(fill = "gray91", color = "gray91"),
        panel.border = element_rect(fill = NA, color = "gray91", size = 2)) + 
  labs(title = "Downed woody debris, by fuel class", 
       caption = "Note: y-axis scale differs by fuel class", 
       y = "Mean tons per acre",  
       x = "Year",
       # Rename legend title
       fill = "Year") + 
  # Reduce font size for y-axis labels 
  # font("y.text", size = 9) +
  facet_wrap(~lab_fuel,  scales = "free", nrow = 1)

p_class

```

We observed a decrease in fuel load immediately after the fire (2017; 0-y post-fire) for all classes of downed woody debris except 1000-hr sound. Specifically, the fuel load in 2017 for 1-hr, 10-hr, 100-hr, and 1000-hr rotten classes was much lower than pre-fire values; variance was unsurprising. Further, mean values for these four classes remained lower than 2016 values for all post-fire years.

The mean for 1000-hr sound fuels increased immediately after the fire and remained greater than 2016 values for all post-fire years. Variance for this fuel class was elevated in 2017. The temporal trend for all downed woody debris was more similar to that for the 1000-hr sound fuel class, and showed less similarity to those for other classes. These results suggest that 1000-hr sound fuel was an influential driver of the post-fire patterns observed for all downed woody debris.

### Check assumptions

##### *Check for outliers*

There were nine extreme outliers in the data set. Most outliers (n = 6) were from the year 2021. The 1-hr and 100-hr classes each had 4 outliers.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
outliers_class <- 
  input_class %>%
  group_by(year, fuel_class) %>%
  identify_outliers(value)  %>%
  mutate(data_type = index_type) %>%
  clean_names() %>%
  select(data_type, fuel_class, year, plot_id, value, 
         starts_with("is")) %>%
  filter(is_extreme == TRUE)

outliers_class %>%
  fxn_kable() 

```

##### *Check for normality*

Note: The data for the 1000-hr sound fuel class were excluded from the normality test; an abundance of zeros in this subset caused the Shapiro test to fail. However, the data for the 1000-hr sound fuel class are shown in a QQ plot below.

Half of the fuel class year combinations were not normally distributed, as assessed by Shapiro-Wilk's test (n = 8 of 16 combinations; hr1000s is excluded). Four combinations included the 1000-hr rotten fuel class; three included year 2017.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
norm_class <- 
  input_class %>%
  filter(fuel_class %nin% "hr1000s") %>%
  select(id = plot_id, 
         time = year, 
         score = value, 
         treatment = fuel_class) %>%
  group_by(time, treatment) %>%
  shapiro_test(score)  %>%
    clean_names() %>%
  mutate(data_type = index_type,
         statistic = fxn_digit(statistic), 
         is_normal = p>0.05) %>%
  select(data_type, 
         fuel_class = treatment, 
         year = time, 
         is_normal,
         p, 
         statistic) %>%
  arrange(is_normal, fuel_class, year) %>%
    filter(is_normal == FALSE)

norm_class %>%
     fxn_kable()

# norm_class %>%
#   group_by(fuel_class, year) %>%
#   count() %>%
#   ungroup() %>%
#   spread(fuel_class, n) 


```

Overall, the distribution of values in the QQ plot didn't look terrible. As expected, the values for 1000-hr sound strayed from the reference line; this was consistent with a distribution that violated the assumptions of the Shapiro test.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggqqplot(input_class, "value",
         facet.by = "fuel_class",
         color = "year",
         palette = colors_year)

```

##### *Normalize data and repeat checks*

We subset the plot-level values for downed woody debris by fuel class, then applied an ordered quantile transformation to normalize the plot-level values for total downed woody debris, calculated as:

g(x) = psi\^-1 \* ((rank(x) - .5) / (length(x)))

Where psi refers to the standard normal cumulative distribution function, rank(x) refers to each observation's rank, and length(x) refers to the number of observations. The ordered quantile transformation is a rank-based procedure by which the values of a vector are mapped to their percentile, which is then mapped to the same percentile of the normal distribution. Without the presence of ties, this essentially guarantees that the transformation leads to a uniform distribution. Values were standardized by fuel class upon normalization to have a mean of 0 and standard deviation of 1.

The following table lists the transformation applied to each fuel class, as well as the mean and standard deviation (sd) for raw and transformed values. Note the untransformed mean for 1000-hr rotten is \>8, a much larger value compared to those of the other classes that range from 0.162 to 1.016.

```{r, echo=FALSE, warning=FALSE}
# List the transformations to use by fuel class
# list_transform <- c("log_x", "log_x", "log_x", "arcsinh_x", "arcsinh_x")

# 1-hr: ordernorm
# 10-hr: ordernorm, sqrt
# 100-hr: nothing looks great, maybe ordernorm?
# 1000-hr rotten: nothing looks great, maybe sqrt? maybe ordernorm?
# 1000-hr sound: raw values look good, then ordernorm  

# # Apply functions to normalize and standardize the data by fuel class 
# bn_hr_0001 <- fxn_transform_log(index_class = "hr0001")
# bn_hr_0010 <- fxn_transform_log(index_class = "hr0010")
# bn_hr_0100 <- fxn_transform_log(index_class = "hr0100")
# bn_hr_1000r <- fxn_transform_arcsine(index_class = "hr1000r")
# bn_hr_1000s <- fxn_transform_arcsine(index_class = "hr1000s")

# Combine subsets 
input_class_transform <- fxn_tranform_ordnorm() %>%
  rename(value = value_ordnorm)

input_class_transform %>%
  select(fuel_class, value_raw, value) %>% 
  gather(metric, value, value_raw:value) %>%
  group_by(fuel_class, metric) %>%
  summarize(mean = fxn_digit(mean(value, na.rm = TRUE)), 
            sd = fxn_digit(sd(value, na.rm = TRUE))) %>%
  ungroup() %>%
  gather(statistic, number, mean:sd) %>%
  spread(fuel_class, number) %>%
  fxn_kable

```

The following plot shows the raw (untransformed) values on the left and the transformed values on the right.

```{r, echo=FALSE, warning=FALSE}
input_class_transform %>%
    rename(raw_values = value_raw, transformed_values = value) %>%

  gather(metric, value, raw_values:transformed_values) %>%
  ggdensity(x = "value", color = "fuel_class") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_grid(metric ~ fuel_class, scales = "free") +
  theme(legend.position = "right", 
        axis.text = element_text(size = 8),
        panel.spacing = unit(1, "lines"),
        legend.text=element_text(size=8),
        # strip.text = element_text(size = 8),
        strip.background = element_rect(fill = "gray91", color = "gray91"),
        panel.border = element_rect(fill = NA, color = "gray91", size = 2)) + 
  labs(title = "Distribution of values, by fuel class and transformation",
       caption = "Note: axis scales differ between plots",
       x = "Mean tons per acre",
       y = "Density", 
       color = "Fuel class") + 
  xlim(-5, 10)


```

The transformed values for downed woody debris were normally distributed for more of the fuel class x year combinations; but still had 5 that were not normally distributed, as assessed by Shapiro-Wilk's test.

```{r, echo=FALSE}
# Test normality ----
norm_class_transform <- 
  input_class_transform %>%
  filter(fuel_class %nin% "hr1000s") %>%
  select(id = plot_id, 
         time = year, 
         score = value, 
         treatment = fuel_class) %>%
  group_by(time, treatment) %>%
  shapiro_test(score)  %>%
  clean_names() %>%
  mutate(data_type = index_type,
         statistic = fxn_digit(statistic), 
         is_normal = p>0.05) %>%
  select(data_type, 
         fuel_class = treatment, 
         year = time, 
         is_normal,
         p, 
         statistic) %>%
  arrange(is_normal, fuel_class, year) %>%
    filter(is_normal == FALSE) 


norm_class_transform %>%
     fxn_kable()


```

However, the points on the QQ plot were mostly near the reference line, suggesting an approximately normal-ish distribution.

```{r, echo=FALSE}
ggqqplot(input_class_transform, 
         "value", 
         facet.by = "fuel_class", 
         color = "year",
         palette = colors_year) +
  labs(caption = "Values have been normalized and standardized by fuel class")

```

All but one of the nine outliers detected in the untransformed data were absent after the transformation was applied.

```{r, echo=FALSE}

input_class_transform %>%
  group_by(year) %>%
  identify_outliers(value)  %>%
  mutate(data_type = index_type) %>%
  clean_names() %>%
  select(data_type, fuel_class, year, plot_id, value, 
         starts_with("is")) %>%
  filter(is_extreme == TRUE) %>%
  fxn_kable()
    
```

### Repeated measures ANOVA test

There was a statistically significant two-way interaction between downed woody debris class and year, F(12, 96) = 5.302, p-adj. = 2.6e-06, ges = 0.242. A significant two-way interaction indicates that the impact of fuel class on fuel load depends on year (and vice versa).

```{r, echo=FALSE, warning=FALSE}
# Two-way ANOVA: Interaction between year x fuel class  
input_class_transform  %>%
  fxn_aov2_me() %>%
  fxn_signif() %>%
  relocate(data_type, 
             effect, 
             starts_with("p_adj"), 
             statistic, 
             starts_with("d_"), 
             ges) %>%
  fxn_kable()

```

### Post hoc tests

We found a significant main effect of year on fuel load for 1-hr, 10-hr (p \< 0.001), 100-hr (p \< 0.01), and 1000-hr sound (p \< 0.05) fuel classes. No significant effect of year on fuel load was found for the 1000-hr sound class. One thing the pairwise comparison by fuel class had in common: For the classes with a significant main effect of year on fuel load, the fuel load in 2016 was significantly different from that in 2021.

```{r, echo=FALSE, warning=FALSE}
# Main effect of year, by fuel class 

input_class_transform %>%
  group_by(fuel_class) %>%
  fxn_aov_me() %>%
  fxn_signif() %>%
  select(-p, 
         -p_05) %>%
  relocate(data_type, 
             fuel_class, 
             effect, 
             starts_with("p_adj"), 
             statistic, 
             starts_with("d_"), 
             ges) %>%
  fxn_kable()
    

```

The following sections present the results from post hoc tests by fuel class for (1) the main effect of fuel class on fuel load at each time point, and (2) significant pairwise differences between years.

```{r, echo=FALSE}
# Pairwise comparison of year, by fuel class ----
pwc_yr_type <- 
  input_class_transform %>%
  group_by(fuel_class) %>%
  fxn_pwc2() %>%
  relocate(data_type, 
             fuel_class, 
             starts_with("group"), 
             starts_with("p_adj"), 
             statistic, 
             starts_with("d")) %>%
  arrange(fuel_class, p_adj)

# Create annotations to visualize p-values on boxplot ----
# pwc_yr <- 
#   input_class_transform %>%
#   filter(fuel_class %in% c("hr0001", "hr0010")) %>%
#   group_by(fuel_class) %>%
#   pairwise_t_test(value ~ year, paired = TRUE, p.adjust.method = "bonferroni")  %>%
#   add_xy_position(x = "fuel_class") %>%
#   filter(fuel_class %nin% c("hr1000r", "hr1000s"))

# Create boxplot comparing both fuel classes by year  ----
# input_class_transform %>%
#   mutate(yr_2digit = str_sub(year, 3, 4)) %>%
#   # filter(fuel_class %in% c("hr0001", "hr0010")) %>%
#    ggboxplot(x = "yr_2digit", 
#             y = "value_raw", 
#             fill = "lab_year_abbr", 
#             palette = colors_year, 
#             outlier.size = 0.3,
#             width = 0.6) +
#   # ylab("Mean tons per acre") + 
#   # xlab()
#   labs(title = "Downed woody debris, by class and year", 
#        fill = "Year", 
#        caption = "Note: y-axis scale differs by fuel class", 
#        x = "Year", 
#        y = "Mean tons per acre") +
#   theme(legend.position = "top", 
#         # axis.title.x = element_blank(),
#         # axis.text.x = element_blank(),
#         axis.text = element_text(size = 10),
#         panel.spacing = unit(1, "lines"))  +
#   facet_wrap("lab_fuel", scales = "free", nrow = 1)  
#   
# stat_pvalue_manual(pwc_yr, 
  #                    x = "year", 
  #                    label = "p.adj.signif",
  #                    tip.length = 0, 
  #                    hide.ns = TRUE) 



```

#### *1-hour vs. Year*

A post hoc pairwise comparison showed a significant difference in 1-hr fuel load between 2016-2017, 2016-2021, and 2017-2019 (p \< 0.01); as well as 2019-2021 (p \< 0.05). No other comparisons were significant.

```{r, echo=FALSE, warning=FALSE}
fxn_posthoc_plot_bt(index_subset = "hr0001", input_class_transform)  
fxn_posthoc_table(index_subset = "hr0001", input_class_transform)

```

#### *10-hour vs. Year*

A post hoc pairwise comparison showed a significant difference in 10-hr fuel load between 2016-2017 (p \< 0.001), 2016-2021 (p \< 0.01), and 2016-2019 (p \< 0.05). No other comparisons were significant.

```{r, echo=FALSE, warning=FALSE}

fxn_posthoc_plot_bt(index_subset = "hr0010", input_class_transform)
fxn_posthoc_table(index_subset = "hr0010", input_class_transform)

```

#### *100-hour vs. Year*

A post hoc pairwise comparison showed no significant difference in 100-hr fuel load between years 2016-2021 at a significance level of \< 0.05.

```{r, echo=FALSE, warning=FALSE}

fxn_posthoc_plot_bt(index_subset = "hr0100", input_class_transform)
fxn_posthoc_table(index_subset = "hr0100", input_class_transform)

```

#### *1000-hour rotten vs. Year*

Post hoc comparisons showed no significant difference in 1000-hr rotten fuel load between years at a significance level of \< 0.05.

```{r, echo=FALSE, warning=FALSE}
fxn_posthoc_plot_bt(index_subset = "hr1000r", input_class_transform)
fxn_posthoc_table(index_subset = "hr1000r", input_class_transform)


```

#### *1000-hour sound vs. Year*

Post hoc comparisons showed no significant difference in 1000-hr sound fuel load between years at a significance level of \< 0.05.

Note: The comparison lines in the plot for 1000-hour sound do not indicate significance. They are a flaw caused by NaN values in the significance test result.

```{r, echo=FALSE, warning=FALSE}
fxn_posthoc_plot_bt(index_subset = "hr1000s", input_class_transform)
fxn_posthoc_table(index_subset = "hr1000s", input_class_transform)


```

------------------------------------------------------------------------

```{r, echo=FALSE}
tibble("Date of last revision" = Sys.Date()) %>%
    knitr::kable()
```
