---
title: 'Coarse woody debris: Pre- and post-thinning'
author: "Morgan Gray"
date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)

# Load libraries
rm(list = ls())
library(tidyverse)   ## To manipulate data frames
library(here)   ## To manage directories
library(janitor)   ## To clean data frames
library(rstatix)  ## For repeated measure ANOVA
library(broom)  ## To format output tables from statistical tests
# library(lemon)  ## To manipulate faceted ggplots
library(kableExtra)  ## For tables in rmarkdown
library(ggpubr)  ## For qq plots
library(patchwork)  ## To arrange faceted plots

# Define file paths 
path_r <- here("R")
path_fxn <- here(path_r, "functions")
path_lookup <- here("input/lookup-tables")
path_derived <- here("input/data_3-derived")

# Define plot colors 
colors_thin_bright <- c("#9d9596", "#069879")
colors_thin_faded <- c("#bfbabb", "#75bca8")

# Source functions 
source(file = here(path_fxn, "basic-functions.R"))
source(file = here(path_fxn, "statistical-tests.R"))
source(file = here(path_fxn, "data-transformation.R"))
# source(file = here(path_fxn, "plot-themes.R"))

```

### Read and summarize derived data (plot-level total, plot-level mean)

```{r raw_read}
# Read data tables for derived total and mean, with metric units 
input_wd <- 
  read_csv(here(path_derived, "thin_total-by-plot-type-trmt.csv")) %>%
  bind_rows(read_csv(here(path_derived, "thin_mean-by-plot-type-class-trmt.csv"))) %>%
  arrange(survey, plot_id, data_type, fuel_class) %>%
  mutate_if(is.character, as_factor)  %>%
  mutate(value = fxn_digit(value_si), 
         timing = ifelse(survey %in% "cont", "survey1", "survey2")) %>%
  select(-value_si, 
         -units) %>%
  rename(units = units_si, 
         metric = statistic) %>%
  relocate(c(metric, subset), .after = value)  %>%
  filter(data_type %in% "wd")

# Create subset for mean by fuel class
wd_class <-
  input_wd %>%
  filter(data_type %in% "wd", 
         metric %in% "mean")

# Create subset for total (all fuel classes combined)
wd_total <-
  input_wd %>%
  filter(data_type %in% "wd", 
         metric %in% "total") %>%
  remove_empty("cols")

# Define helpers 
index_wd <- "Coarse woody debris"
list_classes_wd <- unique(input_wd$fuel_class)
index_units_lab_wd <- "Mean fuel load (metric tons per hectare)"

```

```{r raw_means}
# Pre- and post-thinning mean by fuel class 
input_wd %>%
  group_by(lab_type, lab_fuel, lab_thin, metric, units) %>% 
  summarize(mean = mean(value)) %>%
  spread(lab_thin, mean) %>%
  fxn_kable()
```

```{r raw_zeros}
# 1000-hr rotten (hr1000r) is not normally distributed
#   29 of the 30 transect measurements were 0 
#   9 of 10 plot-level means were 0 (pre-thin: 4/5, post-thin: 5/5)
input_wd %>%
  filter(value == 0) %>%
  group_by(lab_type, lab_fuel, lab_thin) %>% 
  count() %>%
  fxn_kable()
```

### Check assumptions

Subset data subset by fuel class and timing because these are the groupings that will be evaluated with statistical tests.

```{r raw_outliers}
# Determine if there are extreme outliers 
#
# No extreme outliers for total CWD
# Outlier for total pre-thinning in FOR08
#
# Three extreme outliers by fuel class : 
#   Pre-thinning hr0100 and hr1000r in FOR08
#   Post-thinning hr0001 in FOR06
# 
# Overall it looks like something was different about FOR08 during pre- and post-thinning surveys
#
input_wd %>%
  group_by(lab_type, lab_fuel, lab_thin, metric, units) %>% 
  identify_outliers(value) %>%
  clean_names() %>%
  select(lab_type, 
         lab_fuel, 
         lab_thin, 
         plot_id,
         value,
         metric, 
         units,
         starts_with("is")) %>%
  arrange(desc(is_extreme), metric, lab_fuel) %>%
  # filter(is_extreme == TRUE) %>%
  fxn_kable()
```

```{r raw_normality}
# Evaluate normality 
#
# Shapiro test results indicated total CWD was normally distributed at each time point
#
# Shapiro test results by fuel class indicated 100-hr, 1000-hr sound (pre-thin) were not normally distributed 
# 
input_wd %>%
  # Exclude hr1000r because we already know it's not normal
  filter(fuel_class %nin% "hr1000r") %>%
  # Rename columns for shapiro test
  select(id = plot_id, 
         time = lab_thin, 
         score = value, 
         lab_type,
         lab_fuel, 
         metric) %>%
  group_by(lab_type, lab_fuel, time, metric) %>%
  shapiro_test(score) %>%
  clean_names() %>%
  mutate(statistic = fxn_digit(statistic), 
         # p = fxn_digit(p),
         is_normal = p>0.05) %>%
  select(lab_type, 
         lab_fuel, 
         lab_thin = time, 
         metric,
         is_normal,
         p, 
         statistic) %>%
  arrange(is_normal, metric, lab_fuel) %>%
  filter(is_normal == FALSE) %>%
  fxn_kable()
```

```{r raw_qq_plot}
# Create QQ plots for raw data subset by fuel_class and treatment   
ggqqplot(input_wd, 
         "value", 
         palette = colors_thin_bright,
         color = "lab_thin") + 
  facet_wrap(~lab_fuel, 
             scales = "free") + 
  theme(legend.position = "top") +
  geom_hline(yintercept = 0, 
             linetype = "longdash", 
             color = "gray5")  
```

### Identify appropriate transformation
Apply a series of transformations to each subset, then evaluate impact on outliers and normality.

```{r transformation_eval}
# Use four normalization functions from the bestNormalize package to create transformed subsets:
#   arcsinh_x
#   log_x
#   orderNorm
#   sqrt_x
# 
# Most important for subsets that had extreme outliers or were not normally distributed:
# Outlier for total: 
#   Pre-thin in FOR08
#
# Extreme outliers by fuel class: 
#   Pre-thin hr0100 in FOR08
#   Pre-thin hr1000r in FOR08
#   Post-thin hr0001 in FOR06
#
# Not normally distributed:
#   Pre-thin hr0100  
#   Pre-thin hr1000s
#
wd_transform_eval <-
  fxn_tranform_eval(index_data = input_wd,
                    index_list = list_classes_wd)  %>%
  gather(transform, value, value_raw:value_sqrt)

```

```{r transformation_outliers}
# Check outliers for each transformation  
check_transform_outlier <-
  wd_transform_eval %>%
  group_by(timing, 
           fuel_class, 
           transform) %>%
  identify_outliers(value) %>%
  clean_names() %>%
  select(transform,
         fuel_class,
         timing,
         plot_id,
         value,
         starts_with("is")) %>%
  arrange(desc(is_extreme), transform, fuel_class) 

# Pre-thin total in FOR08: No extreme outliers for orderNorm, log_x, arcsinh_x, sqrt_x
check_transform_outlier %>%
  filter(fuel_class %in% "all" & timing %in% "survey1" & plot_id %in% "FOR08") %>%
  fxn_kable()

# Pre-thin hr0100 in FOR08: No extreme outliers for orderNorm, log_x, arcsinh_x
check_transform_outlier %>%
  filter(fuel_class %in% "hr0100" & timing %in% "survey1" & plot_id %in% "FOR08") %>%
  fxn_kable()
       
# Post-thin hr0001 in FOR06: No extreme outliers for orderNorm 
check_transform_outlier %>%
  filter(fuel_class %in% "hr0001" & timing %in% "survey2" & plot_id %in% "FOR06") %>%
  fxn_kable()

# Pre-thin hr1000r in FOR08: No improvement with any transformation 
check_transform_outlier %>%
  filter(fuel_class %in% "hr1000r" & timing %in% "survey1" & plot_id %in% "FOR08") %>%
  fxn_kable()

```

```{r transformation_normality}
# Check normality for each transformation 
check_transform_normal <-
  wd_transform_eval %>%
  # Exclude hr1000r because we already know it's not normal
  filter(fuel_class %nin% "hr1000r") %>%
  # Rename columns for shapiro test
  select(id = plot_id,
         time = timing,
         score = value,
         fuel_class, 
         transform) %>%
  group_by(fuel_class, 
           time, 
           transform) %>%
  shapiro_test(score) %>%
  clean_names() %>%
  mutate(statistic = fxn_digit(statistic), 
         is_normal = p>0.05) %>%
  select(fuel_class,
         timing = time,
         transform,
         is_normal,
         p,
         statistic)

# Pre-thin hr0100: Normal distribution with orderNorm, log_x, arcsinh_x 
check_transform_normal %>%
  filter(fuel_class %in% "hr0100" & timing %in% "survey1", 
         is_normal == TRUE) %>%
  fxn_kable()

# Pre-thin hr1000s: No transformation normalized hr1000s
check_transform_normal %>%
  filter(fuel_class %in% "hr1000s" & timing %in% "survey1",
         is_normal == TRUE) %>%
  fxn_kable()

```

```{r transformation_density_plot}
# The orderNorm transformation looks pretty good for the total and mean subsets
#
wd_transform_eval %>%
  filter(transform %nin% "value_sqrt") %>%
  ggdensity(x = "value", 
            color = "fuel_class") +
  geom_vline(xintercept = 0, 
             linetype = "dashed") +
  facet_grid(fuel_class~transform, 
             scales = "free") +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "right") +
  labs(title = "Distribution of values, by fuel class and transformation",
       caption = "Note: axis scales differ between plots",
       x = "Mean tons per acre",
       y = "Density") +
  xlim(-5, 10)
```

```{r transformation_qq_plot}
# The orderNorm transformed subsets look ok compared to the raw or standardized data
wd_transform_eval %>%
  filter(transform %in% c("value_raw", "value_std",  "value_ordnorm")) %>%
  mutate(transform = str_remove_all(transform, "value_")) %>%
  ggqqplot("value", 
           palette = colors_thin_bright,
           color = "timing") + 
  # To let scales on y-axis vary between faceted columns
  ggh4x::facet_grid2(transform ~ fuel_class, 
                     scales="free", 
                     independent = "all") +
  theme(legend.position = "top", 
        axis.text = element_text(size = 9)) +
  geom_hline(yintercept = 0, 
             linetype = "longdash", 
             color = "gray5")  +
  scale_x_continuous(breaks = c(-1, 0, 1))

```

### Create transformed data as input for statistical tests

```{r transformed_read}
wd_transform <-
  read_csv(here(path_derived, "thin_wd_transformed_metric-units.csv")) %>%
  arrange(timing, plot_id, data_type, fuel_class) %>%
  # Non-numeric variables must be factors for rstatix 
  mutate_if(is.character, as_factor)

wd_transform_total <-
  wd_transform %>%
  filter(data_type %in% "wd",
         metric %in% "total")

wd_transform_class <-
  wd_transform %>%
  filter(data_type %in% "wd",
         metric %in% "mean")

```


### Evaluate total fuel load by treatment 

```{r total_boxplot}
# Visualize the plot-level total amount of coarse woody debris by treatment 
# Use raw data here (instead of transformed values) for easier interpretation by viewer
# A boxplot by treatment showed a post-thinning increase in depth
p_total <- 
  wd_transform_total %>%
  ggboxplot(x = "lab_thin", 
            y = "value_raw", 
            fill = "timing", 
            outlier.size = 0.8,
            palette = colors_thin_faded) + 
  geom_hline(yintercept = 0, 
             linetype = "longdash", 
             color = "gray30") +
  theme(legend.position = "none", 
        axis.title.x = element_blank()) + 
  labs(title = paste0(index_wd, " (all)"), 
       y = index_units_lab_wd) 

p_total
```

```{r total_main_effect}
# Test for an effect of treatment on total CWD load (all CWD fuel classes combined)
# No significant difference in fuel load between pre-thin and post-thin values 
#
wd_transform_total %>%
  # Use transformed values as input for statistical tests
  fxn_aov_me(index_value = "value_tran", 
             index_id = "plot_id", 
             index_time = "lab_thin")  %>%
  fxn_signif_adj() %>%
  mutate(data_type = index_wd,
         fuel_class = "All") %>%
  select(data_type, 
         fuel_class, 
         effect, 
         starts_with("p_adj"),
         statistic, 
         starts_with("d_"), 
         ges) %>%
  fxn_kable()

```

### Evaluate mean fuel load by fuel class and treatment 
```{r class_boxplot}
# Visualize the plot-level mean amount of coarse woody debris by fuel class and treatment 
# Use raw data here (instead of transformed values) for easier interpretation by viewer
# A boxplot by treatment showed a post-thinning increase in depth for all fuel classes except 1000-hr rotten (which had very sparse data)
p_class <-
  wd_transform_class %>%
  ggboxplot(x = "lab_thin", 
            y = "value_raw", 
            fill = "lab_thin", 
            outlier.size = 0.8,
            palette = colors_thin_faded) + 
  geom_hline(yintercept = 0, 
             linetype = "longdash", 
             color = "gray30") +
  theme(legend.position = "top", 
        axis.title.x = element_blank(), 
        axis.text.x = element_blank()) + 
  labs(title = paste0(index_wd, " (by fuel class)"), 
       y = index_units_lab_wd, 
       fill = "Timing") +
  facet_wrap(~lab_fuel, nrow = 1, scales = "free") 

p_class
```
### START WORKING HERE
```{r class_interaction}
# Test for an interaction between treatment and fuel class on mean CWD load 
# wd_transform_class  %>%
#   filter(fuel_class %nin% "all") %>%
#   fxn_aov2_me(index_value = "value_tran", 
#               index_id = "plot_id", 
#               index_time = "timing", 
#               index_variable = "fuel_class") %>%
#   fxn_signif_adj() %>%
#   mutate(data_type = index_wd) %>%
#   select(data_type, 
#            effect, 
#              starts_with("p_adj"), 
#              statistic, 
#              starts_with("d_"), 
#              ges) %>%
#   fxn_kable()
# Significant interaction between treatment and fuel class 
```

```{r class_pairwise}
# Pairwise comparisons in fuel load between pre- and post-thin values, by fuel class
# Effect of treatment on each fuel class 
# wd_transform_class %>%
#     fxn_pwc2(index_value = "value_tran", 
#             index_id = "plot_id", 
#             index_time = "lab_thin", 
#             index_variable = "fuel_class") %>%
#     select(fuel_class = variable, 
#              starts_with("group"), 
#              starts_with("p_adj"), 
#              statistic, 
#              starts_with("d")) %>%
#     fxn_kable()
# 
# litter <- 
#   fxn_posthoc_plot_bt(index_input = wd_transform_class,
#                     index_subset <- "litter",
#                     index_time <- "lab_thin",
#                     index_value <- "value_tran") +
#   scale_fill_manual(values = colors_thin_faded) + 
#   theme(legend.position = "none")
# 
# duff <- 
#   fxn_posthoc_plot_bt(index_input = wd_transform_class,
#                     index_subset <- "duff",
#                     index_time <- "lab_thin",
#                     index_value <- "value_tran") +
#   scale_fill_manual(values = colors_thin_faded)  + 
#   theme(axis.title.y = element_blank(), 
#         legend.position = "none") + 
#   labs(caption = "Values have been transformed and standardized by fuel class")
# 
# 
# litter + duff
```

