---
title: 'Fuel analysis for Oak Symposium: Duff and litter'
author: "Morgan Gray"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries -----
rm(list = ls())
library(tidyverse)   ## To manipulate data frames
library(here)   ## To manage directories
library(janitor)   ## To clean data frames
library(rstatix)  ## For repeated measure ANOVA
library(broom)  ## To format output tables from statistical tests
library(lemon)  ## To manipulate faceted ggplots
library(kableExtra)  ## For tables in rmarkdown
library(ggpubr)  ## For qq plots

# Define file paths ----
path_r <- here("R")
path_fxn <- here(path_r, "functions")
path_lookup <- here("input/lookup-tables")
path_raw <- here("input/data_1-raw")
path_tidy <- here("input/data_2-tidy")
path_derived <- here("input/data_3-derived")

# Define helpers ----
index_type <- "Duff & litter"
index_units_si <- "Depth in centimeters"

# Define plot colors ----
plot_colors <- 
  read_csv(here(path_lookup, 
                "plot-colors.csv")) %>%
  arrange(palette, palette_subset, levels)

colors_year <- 
  plot_colors %>%
  filter(palette %in% "year", 
         palette_subset %in% "1pre_3post") %>%
  pull(hex_code) 

colors_dl <- c("#7da8b0", "#c4c4c4")

# Source functions ----
source(file = here(path_fxn, "basic-functions.R"))
source(file = here(path_fxn, "statistical-tests.R"))
source(file = here(path_fxn, "data-transformation.R"))

```

------------------------------------------------------------------------

## Overview

The aim of this research was to understand the impact of of the 2017 Tubbs fire on fuel loads. The objective of this research was to evaluate the impact of the 2017 Tubbs fire on duff and litter fuel loads over time. To meet this objective, we compared the pre- and post-fire levels of duff and litter at nine long-term vegetation monitoring plots. Surveys were conducted at four time points: 2016 (1y pre-fire), 2017 (0y post-fire), 2019 (2y post-fire), and 2021 (4y post-fire). We conducted surveys along two transects per plot; each transect had two quadrats. As a result, each time point had 36 data points: 9 plots x 2 transects x 2 quadrats.

Using the duff and litter data set as input, we asked the following questions:

1.  Did the total fuel load (all duff and litter combined) differ between years?
2.  Did fuel load differ between years for duff and/or litter? If so, which years were different?

#### Duff and litter data

We answered these questions using a time series of plot-level data for (1) the total amount of duff and litter and (2) the mean amount of each fuel class (duff, litter). We defined "total (or all) duff and litter" as the combined total of the duff and litter within a plot.

First, we calculated the plot-level mean (n = 4 quadrats) for each fuel class (duff, litter) at each time point. A subset of the data frame with the plot-level mean for each fuel class by year is shown below.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
input_class <- 
  read_csv(here(path_derived, "tubbs_mean-by-plot-type-class-yr.csv")) %>%
  arrange(year) %>%
  mutate(year = as_factor(year), 
         lab_year = as_factor(lab_year), 
         lab_year_abbr = as_factor(lab_year_abbr), 
         value = fxn_digit(value_si), 
         statistic = "mean",
         units = index_units_si) %>%
  relocate(c(statistic, subset), .after = value)  %>%
  filter(lab_type %in% index_type)

head(input_class) %>%
  select(fuel_type = lab_type, 
         fuel_class = lab_fuel, 
         plot_id, 
         year,
         value, 
         units,
         statistic) %>%
  fxn_kable()  

```

Next, we summed the means for the duff and litter fuel classes to find the plot-level total at each time point. Below is a subset of the data frame for the plot-level values of duff and litter by year.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
input_total <-
  read_csv(here(path_derived, "tubbs_total-by-plot-type-yr.csv")) %>%
  arrange(year) %>%
  mutate(year = as_factor(year), 
         lab_year = as_factor(lab_year), 
         lab_year_abbr = as_factor(lab_year_abbr), 
         lab_fuel = "All", 
         value = fxn_digit(value_si), 
         statistic = "total", 
         units = index_units_si) %>%
  relocate(c(statistic, subset), .after = value)  %>%
  filter(lab_type %in% index_type)

head(input_total) %>%
  select(fuel_type = lab_type, 
         fuel_class = lab_fuel, 
         plot_id, 
         year,
         value, 
         units,
         statistic) %>%
    fxn_kable()

```

#### Statistical methods

Statistically significant differences were identified using repeated measures analysis of variance (ANOVA) and post hoc comparisons. P-values were adjusted using the Bonferroni multiple testing correction method.

The repeated-measures ANOVA is used for analyzing data where same subjects are measured more than once. This test is also referred to as a within-subjects ANOVA or ANOVA with repeated measures. The "within-subjects" term means that the same individuals (here, individuals are plots) are measured on the same outcome variable under different time points. The main goal of a repeated measures ANOVA is to evaluate if there is a statistically significant interaction effect between within-subjects factors in explaining a continuous outcome variable. The repeated measures ANOVA makes the following assumptions about the data:

-   No significant outliers in any cell of the design
-   Normality: the outcome (or dependent) variable should be approximately normally distributed in each cell of the design
-   Assumption of sphericity: the variance of the differences between groups should be equal

We assessed outliers using the the interquartile range (IQR; IQR = Q3 - Q1). Values above Q3 + 1.5xIQR or below Q1 - 1.5xIQR are considered as outliers. Values above Q3 + 3xIQR or below Q1 - 3xIQR are considered as extreme points (or extreme outliers). Q1 and Q3 are the first and third quartile, respectively. Extreme outliers can be due to data entry errors, measurement errors, or unusual values. The outlier may be included if one believes the result will not be substantially affected; this can be evaluated by comparing the result of the ANOVA with and without the outlier.

We assessed normality by visual inspection of a QQ plot for each time point. A QQ plot draws the correlation between a given data and the normal distribution. We also conducted the Shapiro-Wilk test for each time point. Using this method, normally distributed data will have p-value \>0.05.

The assumption of sphericity was checked during the computation of the ANOVA test using the R function anova_test() [rstatix package]. The Mauchly's test was internally used to assess the sphericity assumption, and the Greenhouse-Geisser sphericity correction was automatically applied to factors violating the sphericity assumption.

------------------------------------------------------------------------

## Q1: Did total fuel load differ between years?

The first question we asked was whether there was a significant main effect of year on total fuel load. A one-way repeated measures ANOVA was used to determine whether fuel load was significantly different between the four time points.

### Summary statistics for the plot-level total amount of duff and litter by year

The following table summarizes the plot-level total amount of duff and litter for each time point.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
input_total %>%
  group_by(year, units) %>%
  get_summary_stats(value, type = "mean_sd")  %>%
  mutate(fuel_type = index_type,
         fuel_class = "All",) %>%
  select(fuel_type, 
         fuel_class, 
         year,
         n,
         mean, 
         sd,
         units) %>%
     fxn_kable()

```

### Visualization of the plot-level total amount of duff and litter by year

A box plot of total duff and litter by year showed lower values in post-fire years. Specifically, we observed a decrease in the depth of all duff and litter immediately following the fire (2017; 0-y post-fire). The mean depth of duff and litter did not return to pre-fire levels in subsequent years.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p_total <- 
  input_total %>%
  ggboxplot(x = "lab_year", 
            y = "value", 
            fill = "year", 
            palette = colors_year) + 
  theme(legend.position = "none", 
        axis.title.x = element_blank()) + 
  scale_x_discrete(labels = function(lab_year) str_wrap(lab_year, width = 20)) + 
  labs(title = paste0(index_type, " (all)"), 
       y = paste0(index_units_si, " (mean)"))

p_total

```

### Check assumptions

##### *Check for outliers*

There was one extreme outlier in the data set. Specifically, the value for plot RxF03 in 2017 was 1.738.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
input_total %>%
  group_by(year) %>%
  identify_outliers(value)  %>%
  mutate(data_type = index_type,
         fuel_class = "All") %>%
  clean_names() %>%
  select(data_type, fuel_class, year, plot_id, value, 
         starts_with("is")) %>%
  filter(is_extreme == TRUE) %>%
  fxn_kable()

```

##### *Check for normality*

Plot-level duff and litter load was normally distributed at each time point (p \> 0.01), except for year 2017, as assessed by Shapiro-Wilk's test.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
input_total %>%
  select(id = plot_id, 
         time = year, 
         score = value) %>%
  group_by(time) %>%
  shapiro_test(score)  %>%
    clean_names() %>%
  mutate(data_type = index_type,
         fuel_class = "All", 
         statistic = fxn_digit(statistic), 
         is_normal = p>0.05) %>%
  select(data_type, 
         fuel_class, 
         year = time, 
         is_normal,
         p, 
         statistic) %>%
  arrange(is_normal, year) %>%
      filter(is_normal == FALSE) %>%
  fxn_kable()

```

The QQ plot showed one point in 2017 that was distant from the reference line.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggqqplot(input_total, "value", 
         facet.by = "year", 
         color = "year",
         palette = colors_year)

```

### Repeated measures ANOVA test

Total duff and litter load differed between years. We found a significant main effect of year on duff and litter load (p-adjusted \< 0.001). The mean fuel load for all duff and litter (duff and litter combined) was significantly different between the four years; F(3, 24) = 14.029, p-adjusted (p-adj.) = 1.73e-05, ges = 0.59. The generalized effect size (ges) is the amount of variability due to the within-subjects factor.

```{r, echo=FALSE, warning=FALSE}
# Effect of year (all fuel classes)
input_total %>%
  fxn_aov_me() %>%
  mutate(data_type = index_type,
         fuel_class = "All") %>%
  clean_names() %>%
  select(-p, 
         -p_05) %>%
  relocate(data_type, 
           fuel_class,
           effect, 
           starts_with("p_adj"), 
           statistic, 
           starts_with("d_"), 
           ges) %>%
  fxn_kable()
```

### Post hoc tests

We conducted post hoc pairwise comparisons between the levels of the within-subjects factor (here, year). The result of paired t-tests between years showed a significant difference between pre- and post-fire mean duff and litter levels in 2017 (p-adj. \<0.001) and 2021 (p-adj \<0.01).

The results from the pairwise comparisons are shown below as (1) a boxplot of total duff and litter by year updated with p-values, and (2) a table of test results.

```{r, echo=FALSE}
pwc <- 
  input_total %>%
  pairwise_t_test(value ~ year, paired = TRUE, p.adjust.method = "bonferroni")  %>% 
  add_xy_position(x = "year")

aov_me <- 
  input_total %>%
    anova_test(dv = value, wid = plot_id, within = year)
  
p_total + 
  labs(subtitle = get_test_label(aov_me, detailed = TRUE),
       caption = get_pwc_label(pwc)) +
  stat_pvalue_manual(pwc, hide.ns = TRUE) 

```

```{r, echo=FALSE, warning=FALSE}
# Effect of year (all fuel classes)
input_total %>%
  fxn_pwc() %>%
  mutate(statistic = fxn_digit(statistic),
         data_type = index_type,
         fuel_class = "All") %>%
  select(-p) %>%
  relocate(data_type, 
           fuel_class,
           starts_with("group"), 
           starts_with("p_adj"), 
           statistic, 
           starts_with("d")) %>%
  arrange(p_adj) %>%
  fxn_kable()

```

------------------------------------------------------------------------

## Q2: Did fuel load differ between years by fuel class?

Next, we investigated whether there was a significant change in plot-level fuel load over time when accounting for fuel class (duff, litter). A two-way repeated measures ANOVA was used to determine whether there was a significant interaction between year and fuel class on fuel load.

Here, the effect of year on fuel load was our focal variable of primary concern. However, the effect of year may differ between duff and litter, so fuel_class was considered a moderator variable.

### Summary statistics for duff and litter by year

The following table summarizes duff and litter by class for each time point.

```{r, echo=FALSE}
input_class %>%
  group_by(year, fuel_class, units) %>%
  get_summary_stats(value, type = "mean_sd")  %>%
  arrange(fuel_class, year) %>%
  mutate(fuel_type = index_type) %>%
  select(fuel_type, 
         fuel_class, 
         year,
         n,
         mean, 
         sd,
         units) %>%
  gather(statistic, value, n:sd) %>%
  spread(year, value) %>%
  relocate(units, .after = last_col()) %>%
  fxn_kable()

```

### Visualization of duff and litter by year

A box plot of duff and litter by year showed differing post-fire trends. Although duff and litter both showed a decrease in depth immediately following the fire (2017; 0-y post-fire), the subsequent annual trends differed. The temporal trend for total duff and litter was more similar to that of the litter fuel class, and showed less similarity to that of the duff fuel class. These results suggest that litter was an influential driver of the post-fire patterns observed for duff and litter as a group.

**Duff**\
-- Duff depth remained low in all post-fire years.\
-- Values in 2019 and 2021 were lower than those for 2017.

**Litter**\
-- Litter depth returned to pre-fire levels 2-y after the fire (2019), then showed a second decrease in 2021 (4-y post-fire).\
-- This post-fire increase and decrease in litter depth was similar to the temporal trend observed for duff and litter overall.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p_class <- 
  input_class %>%
  bind_rows(input_total) %>%
  ggboxplot(x = "year", 
            y = "value", 
            fill = "year", 
            palette = colors_year) + 
  theme(legend.position = "none", 
        axis.title.x = element_blank(), 
        panel.spacing = unit(2, "lines"),
        strip.background = element_rect(fill = "gray91",
                                        color = "gray91"),
        panel.border = element_rect(fill = NA,
                                    color = "gray91",
                                    size = 1.5)) + 
  # scale_x_discrete(labels = function(lab_year) str_wrap(lab_year, width = 20)) + 
  labs(title = paste0(index_type), 
       y = paste0(index_units_si, " (mean)")) + 
  facet_wrap(~lab_fuel)

p_class

```

### Check assumptions

##### *Check for outliers*

There were two extreme outliers in the data set. Specifically, the values for duff at plots RxF03 and RxF04 in 2021 were 0.112 and 0.500, respectively.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
input_class %>%
  group_by(year, fuel_class) %>%
  identify_outliers(value)  %>%
  mutate(data_type = index_type) %>%
  clean_names() %>%
  select(data_type, fuel_class, year, plot_id, value, 
         starts_with("is")) %>%
  filter(is_extreme == TRUE) %>%
  fxn_kable()

```

##### *Check for normality*

Plot-level litter and duff values were normally distributed at each time point (p \> 0.01), except for year duff in 2021 and litter in 2017, as assessed by Shapiro-Wilk's test.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
input_class %>%
  select(id = plot_id, 
         time = year, 
         score = value, 
         treatment = fuel_class) %>%
  group_by(time, treatment) %>%
  shapiro_test(score)  %>%
    clean_names() %>%
  mutate(data_type = index_type,
         statistic = fxn_digit(statistic), 
         is_normal = p>0.05) %>%
  select(data_type, 
         fuel_class = treatment, 
         year = time, 
         is_normal,
         p, 
         statistic) %>%
  arrange(is_normal, fuel_class, year) %>%
    filter(is_normal == FALSE) %>%
     fxn_kable()

ggqqplot(input_class, "value", 
         facet.by = "year", 
         color = "fuel_class",
         palette = colors_dl)

```

### Repeated measures ANOVA test

There was a statistically significant two-way interaction between duff and litter and year, F(2, 14) = 10.382, p-adj. = 0.006, ges = 0.329. A significant two-way interaction indicates that the impact of fuel class on fuel load depends on year (and vice versa).

```{r, echo=FALSE, warning=FALSE}
# Two-way ANOVA: Interaction between year x fuel class  
input_class %>%
  fxn_aov2_me() %>%
  fxn_signif_adj() %>%
  # mutate(data_type = index_type) %>%
  relocate(data_type, 
           effect, 
           starts_with("p_adj"), 
           statistic, 
           starts_with("d_"), 
           ges) %>%
  fxn_kable()

```

### Post hoc tests

We found a significant main effect of year on fuel load for both duff and litter. One thing the pairwise comparisons by fuel class had in common: For both duff and litter, the fuel load in 2016 was significantly different from that in 2017 (duff: p \<0.001; litter: p \< 0.01).

```{r, echo=FALSE}
# Create annotations to visualize p-values on boxplot ----
pwc_yr <- 
  input_class %>%
  group_by(fuel_class) %>%
  pairwise_t_test(value ~ year, paired = TRUE, p.adjust.method = "bonferroni")  %>%
  add_xy_position(x = "fuel_class")

# Create boxplot comparing both fuel classes by year  ----
input_class %>%
  ggboxplot(x = "lab_fuel", 
            y = "value", 
            fill = "lab_year_abbr", 
            palette = colors_year, 
            width = 0.6) +
  labs(title = paste0(index_type), 
       y = paste0(index_units_si, " (mean)"),
       fill = "Year") +
  theme(axis.title.x = element_blank(), 
        legend.position = "right")  + 
  stat_pvalue_manual(pwc_yr, tip.length = 0, hide.ns = TRUE) 


```

```{r, echo=FALSE, warning=FALSE}
# Main effect of year, by fuel class 
input_class %>%
  group_by(fuel_class) %>%
  fxn_aov_me() %>%
  fxn_signif_adj() %>%
  select(-p, 
         -p_05) %>%
  relocate(data_type, 
           fuel_class, 
           effect, 
           starts_with("p_adj"), 
           statistic, 
           starts_with("d_"), 
           ges) %>%
  fxn_kable()

```

The following sections present the results from post hoc tests by fuel class for (1) the main effect of fuel class on fuel load at each time point, and (2) significant pairwise differences between years.

##### *Duff vs. Year*

There was a statistically significant main effect of year on duff; F(3, 24) = 62.909, p-adj. = 1.07e-05, ges = 0.853. Pairwise comparisons showed that the mean duff depth was significantly different between all years except 2019/2021. This result indicated that duff depth differed between 2016 (pre-fire) and all post-fire years, as well as between 2017 (0y post-fire) and subsequent years.

```{r, echo=FALSE}
fxn_posthoc_plot(index_subset = "duff", input_class) +
  theme(legend.position = "right")
fxn_posthoc_table(index_subset = "duff", input_class)

 
```

##### *Litter vs. Year*

There was a statistically significant main effect of year on litter; F(1, 10) = 6.83, p-adj. = 0.004, ges = 0.423. Pairwise comparisons showed that the mean litter depth was significantly different between 2016 and 2017 (p-adj. \<0.01), as well as 2017 and 2019 (p-adj. \<0.05).

```{r, echo=FALSE, warning=FALSE}

fxn_posthoc_plot(index_subset = "litter", input_class) +
  theme(legend.position = "right")

fxn_posthoc_table(index_subset = "litter", input_class)

```

##### *Duff vs. Litter*

We found a significant difference in mean depth between duff and litter in 2019 and in 2021.

```{r, echo=FALSE}

# Create annotations to visualize p-values on boxplot ----
pwc_class <- 
  input_class %>%
  group_by(year) %>%
  pairwise_t_test(value ~ fuel_class, paired = TRUE, p.adjust.method = "bonferroni") %>%
  add_xy_position(x = "year")

# Create boxplot comparing fuel class by year  ----
input_class %>%
  ggboxplot(x = "lab_year", 
            y = "value", 
            fill = "lab_fuel", 
            palette = colors_dl, 
            width = 0.6) +
  scale_x_discrete(labels = function(lab_year) str_wrap(lab_year, width = 20)) + 
  labs(title = paste0(index_type), 
       y = paste0(index_units_si, " (mean)"),
       fill = "Fuel class") +
  theme(axis.title.x = element_blank(), 
        legend.position = "right")  + 
  font("x.text", size = 10)  +
  stat_pvalue_manual(pwc_class, tip.length = 0, hide.ns = TRUE) 

# Create summary table  ----
pwc_class %>%
  mutate(data_type = index_type,
         statistic = fxn_digit(statistic), 
         method = "pwc") %>%
  clean_names() %>%
  select(data_type, 
         year,
         group1, group2,
         starts_with("p_adj"), 
         statistic, 
         starts_with("d")) %>%
  fxn_kable()

```

------------------------------------------------------------------------

```{r, echo=FALSE}
tibble("Date of last revision" = Sys.Date()) %>%
    knitr::kable()
```
