---
title: 'Coarse woody debris: Pre- and post-Tubbs fire'
author: "Morgan Gray"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Load libraries -----
rm(list = ls())
library(tidyverse)   ## To manipulate data frames
library(here)   ## To manage directories
library(janitor)   ## To clean data frames
library(rstatix)  ## For repeated measure ANOVA
library(broom)  ## To format output tables from statistical tests
library(kableExtra)  ## For tables in rmarkdown
library(ggpubr)  ## For qq plots
library(bestNormalize)

# Define file paths ----
path_r <- here("R")
path_fxn <- here(path_r, "functions")
path_lookup <- here("input/lookup-tables")
path_derived <- here("input/data_derived")
# Source functions ----
source(file = here(path_fxn, "helpers_woody-debris.R"))
source(file = here(path_fxn, "basic-functions.R"))
source(file = here(path_fxn, "plot-themes.R"))
# ========================================================== -----
# Read input data: derived total and mean ----
input_wd <-
  read_csv(here(path_derived, "tubbs_derived-norm.csv")) %>%
  arrange(time, fuel_type, fuel_class) %>%
  mutate_if(is.character, as_factor)  %>%
  mutate(time = as_factor(time), 
         si_value = fxn_digit(si_value), 
         units = ifelse(fuel_type %in% "wd", index_wd_si_abbr, index_wd_si_abbr)) %>%
  filter(fuel_type %in% "wd")  

wd_class <-
  input_wd %>%
  filter(metric %in% "mean")

wd_total <-
  input_wd %>%
  filter(metric %in% "total")  

```

------------------------------------------------------------------------

## Overview

The aim of this research was to understand the impact of of the 2017 Tubbs fire on fuel loads. Our objective was to evaluate the impact of prescribed fire on fuel loads over time for five coarse woody debris fuel classes: 1-hr, 10-hr, 100-hr, 1000-hr rotten, and 1000-hr sound. To meet this objective, we compared the pre- and post-fire levels of these five lag-time fuel classes at nine long-term vegetation monitoring plots. Surveys were conducted at four time points: 2016 (1y pre-fire), 2017 (0y post-fire), 2019 (2y post-fire), and 2021 (4y post-fire). We conducted surveys along two transects per plot; each transect had one quadrat. Each time point had 18 data points: 9 plots x 2 transects x 1 quadrat.

Using the coarse woody debris data set as input, we asked the following questions:

1.  Did the total fuel load (all fuel classes combined) differ between years?
2.  Did fuel load differ between years by fuel class? If so, which years were different?

#### Coarse woody debris data

We answered these questions using a time series of plot-level data for (1) the total amount of coarse woody debris and (2) the mean amount of each of the five fuel classes. We defined "total (or all) coarse woody debris" as the combined total of the the five fuel classes within a plot: 1-hr, 10-hr, 100-hr, 1000-hr rotten, 1000-hr sound.

We calculated the ***plot-level mean for each fuel class*** to account for replicate surveys within each plot. To calculated the the ***plot-level total*** at each time point, we summed the plot means for all fuel classes by fuel type.

Below is a table that shows the plot-level mean at each time point for coarse woody debris (total and mean by fuel class).

```{r raw_summary}
input_wd %>%
  arrange(plot_id, fuel_class, time) %>%
  select(fuel_class = lab_class, 
         plot_id, 
         time,
         si_value, 
         metric,
         units) %>%
  spread(time, si_value)  %>%
  fxn_kable()
```

#### Statistical methods

Statistically significant differences were identified using repeated measures analysis of variance (ANOVA) and post hoc comparisons. P-values were adjusted using the Bonferroni multiple testing correction method.

The repeated-measures ANOVA is used for analyzing data where same subjects are measured more than once. This test is also referred to as a within-subjects ANOVA or ANOVA with repeated measures. The "within-subjects" term means that the same individuals (here, individuals are plots) are measured on the same outcome variable under different time points. The main goal of a repeated measures ANOVA is to evaluate if there is a statistically significant interaction effect between within-subjects factors in explaining a continuous outcome variable. The repeated measures ANOVA makes the following assumptions about the data:

-   No significant outliers in any cell of the design
-   Normality: the outcome (or dependent) variable should be approximately normally distributed in each cell of the design
-   Assumption of sphericity: the variance of the differences between groups should be equal

We assessed outliers using the the interquartile range (IQR; IQR = Q3 - Q1). Values above Q3 + 1.5xIQR or below Q1 - 1.5xIQR are considered as outliers. Values above Q3 + 3xIQR or below Q1 - 3xIQR are considered as extreme points (or extreme outliers). Q1 and Q3 are the first and third quartile, respectively. Extreme outliers can be due to data entry errors, measurement errors, or unusual values. The outlier may be included if one believes the result will not be substantially affected; this can be evaluated by comparing the result of the ANOVA with and without the outlier.

We assessed normality by visual inspection of a QQ plot for each time point. A QQ plot draws the correlation between a given data and the normal distribution. We also conducted the Shapiro-Wilk test for each time point. Using this method, normally distributed data will have p-value \>0.05.

The assumption of sphericity was checked during the computation of the ANOVA test using the R function anova_test() [rstatix package]. The Mauchly's test was internally used to assess the sphericity assumption, and the Greenhouse-Geisser sphericity correction was automatically applied to factors violating the sphericity assumption.

#### Check assumptions

##### *Check for outliers*

No extreme outliers were found for the total fuels subset.

There were nine extreme outliers in the mean by fuel class data set. Most outliers (n = 6) were from the year 2021. The 1-hr and 100-hr classes each had 4 outliers. We found the following extreme outliers by fuel class:

**1-hour**

-   2021 (RxF03, RxF04, RxF05, RxF08)

**100-hour**

-   2017 (RxF03, RxF08)

-   2021 (RxF04, RxF05)

**1000-hr rotten**

-   excluded here because too many values are close to zero

**1000-hr sound**

-   2016 (RxF04)

```{r raw_outliers}
input_wd %>%
  group_by(time, fuel_class) %>% 
  identify_outliers(si_value) %>%
  clean_names() %>%
  select(fuel_class, 
         # metric,
         time, 
         plot_id, 
         si_value, 
         starts_with("is")) %>%
  filter(is_extreme == TRUE)  %>%
  arrange(fuel_class, time, plot_id) %>%
  fxn_kable()

```

Several points in the QQ plot fell outside the reference range.

Overall, the distribution of values by fuel class in the QQ plot didn't look terrible. As expected, the values for 1000-hr sound strayed from the reference line; this was consistent with a distribution that violated the assumptions of the Shapiro test.

```{r raw_qq_plot}
ggqqplot(input_wd, 
         "si_value", 
         palette = colors_year,
         color = "time") + 
  facet_wrap(~lab_class, scales = "free") + 
  theme(legend.position = "top") +
  geom_hline(yintercept = 0, linetype = "longdash", color = "gray30")

```

##### *Check for normality*
Plot-level values for total fuels were not normally distributed for three of the four time points (2016, 2017, 2019), as assessed by Shapiro-Wilk's test.

The data for the 1000-hr rotten fuel class were excluded from the normality test; an abundance of zeros in this subset caused the Shapiro test to fail. However, the data for the 1000-hr rotten fuel class are shown in a QQ plot below.

Half of the fuel class year combinations were not normally distributed, as assessed by Shapiro-Wilk's test (n = 8 of 16 combinations; hr1000r is excluded). Four combinations included the 1000-hr sound fuel class; three included year 2017.

Distribution not normal for: 
all	2016
all	2017
all	2019

hr0010	2017

hr0100	2017
hr0100	2019
hr0100	2021

hr1000s	2016
hr1000s	2017
hr1000s	2019
hr1000s	2021

```{r raw_normality}
input_wd %>%
  filter(fuel_class %nin% "hr1000r") %>%
  select(id = plot_id, 
         time = time, 
         score = si_value, 
         fuel_class) %>%
  group_by(fuel_class, time) %>%
  shapiro_test(score) %>%
  clean_names() %>%
  mutate(statistic = fxn_digit(statistic), 
         is_normal = p>0.05) %>%
  select(fuel_class, 
         time, 
         is_normal,
         p, 
         statistic) %>%
  arrange(is_normal, fuel_class, time)  %>%
  filter(is_normal == FALSE)  %>%
  fxn_kable()




```

#### Identify the most appropriate transformation
We applied a series of transformations to each subset, then evaluated the resulting data for outliers and normality. To create the transformed subsets, we used scale() from base R and four normalization functions from the bestNormalize package: arcsinh_x, log_x, orderNorm, and sqrt_x.

```{r transformation_eval}
wd_eval <- 
  input_wd %>%
  group_by(lab_class, fuel_class) %>%
  mutate(value_std = as.vector(scale(si_value)), 
         value_arcsine = arcsinh_x(si_value, standardize = TRUE)$x.t, 
         value_log = log_x(si_value, standardize = TRUE)$x.t, 
         value_ordnorm = orderNorm(si_value, standardize = TRUE)$x.t, 
         value_sqrt = sqrt_x(si_value, standardize = TRUE)$x.t) %>%
  ungroup() %>%
  relocate(c(si_value, us_value, value_norm), .after = last_col()) %>%
  gather(method, number, value_std:value_norm) %>%
  select(fuel_class, lab_class, plot_id, time, lab_time_abbr, metric, method, number) 

```

All but one of the nine outliers detected in the untransformed data were absent after the transformation was applied (by fuel class).
Only one outlier was improved by any transformation: 1-hr in 2021 at RxF08, ok with ordnorm.  

```{r transformation_outliers}
wd_eval_outlier <- 
  wd_eval %>%
  group_by(fuel_class, 
           time, 
           method) %>%
  identify_outliers(number) %>%
  clean_names()  %>%
  ungroup() %>%
  select(fuel_class,
         time, 
         method,
         plot_id, 
         number, 
         starts_with("is")) %>%
  filter(is_extreme == FALSE)  

wd_eval_outlier %>%
  mutate(method = str_remove_all(method, "value_")) %>%
  filter(fuel_class %nin% c("all", "hr0010"), 
         time %in% c("2016", "2017", "2021")) %>%
  select(fuel_class, time, plot_id, method, is_extreme) %>%
  spread(method, is_extreme)  %>%
  fxn_kable()

# Most important for subsets that had extreme outliers
# 1-hour
# 2021 (RxF03, RxF04, RxF05, RxF08): RxF08 ok with ordnorm
# 
# 100-hour
# 2017 (RxF03, RxF08): none
# 2021 (RxF04, RxF05): none 
# 
# 1000-hr rotten
# excluded here because too many values are close to zero 
# 
# 1000-hr sound
# 2016 (RxF04): none


```

```{r transformation_qq_plot}
# wd_eval %>%
#   mutate(method = str_remove_all(method, "value_")) %>%
#   # Subset to fuel classes and years with outliers 
#   filter(fuel_class %nin% c("all", "hr0010"),
#          method %in% c("si_value",
#                        "std",
#                        "arcsine",
#                        "log",
#                        "ordnorm", 
#                        "sqrt")) %>%
#   ggqqplot("number", 
#            palette = colors_year,
#            color = "time") + 
#   # To let scales on y-axis vary between faceted columns
#   ggh4x::facet_grid2(lab_class ~ method, 
#                      scales="free", 
#                      independent = "all") +
#   theme(legend.position = "none", 
#         axis.text = element_text(size = 9)) +
#   geom_hline(yintercept = 0, 
#              linetype = "longdash", 
#              color = "gray5")  +
#   scale_x_continuous(breaks = c(-1, 0, 1))

# However, the points on the QQ plot were mostly near the reference line, suggesting an approximately normal-ish distribution.
```

The transformed values for CWD by fuel class were normally distributed for more of the fuel class x year combinations; but still had 5 that were not normally distributed, as assessed by Shapiro-Wilk's test.

```{r transformation_normality}
wd_eval_normal <- 
  wd_eval %>%
  filter(fuel_class %nin% "hr1000r") %>%
  select(id = plot_id, 
         time,
         score = number,
         treatment = fuel_class,
         method) %>%
  group_by(time, treatment, method) %>%
  shapiro_test(score)  %>%
  clean_names() %>%
  mutate(is_normal = p>0.05) %>%
  select(fuel_class = treatment,
         time,
         is_normal,
         method,
         p,
         statistic) %>%
  arrange(is_normal, fuel_class, method, time) %>%
  filter(is_normal == TRUE)

# wd_eval_normal %>%
#   mutate(method = str_remove_all(method, "value_")) %>%
#   filter(fuel_class %nin% c("all", "hr0010"), 
#          time %in% c("2016", "2017", "2021")) %>%
#   select(fuel_class, time, plot_id, method, is_extreme) %>%
#   spread(method, is_extreme)  %>%
#   fxn_kable()

# all	2016	 
# all	2017	 
# all	2019	 

# hr0010	2017 

# hr0100	2017	 
# hr0100	2019	 
# hr0100	2021	 

# hr1000s	2016 
# hr1000s	2017	 
# hr1000s	2019 
# hr1000s	2021 

```

```{r, transformation_density_plot}
# wd_eval %>%
#   mutate(method = str_remove_all(method, "value_")) %>%
#   # Subset to fuel classes and years with outliers 
#   filter(time %in% c("2017", "2021"), 
#          # fuel_class %in% c("all", "duff"),
#          method %in% c("si_value",
#                        # "std",
#                        "arcsine",
#                        "log",
#                        "ordnorm")) %>%
#                        # "sqrt")) %>%
#   ggdensity(x = "number", 
#             color = "time") +
#   geom_vline(xintercept = 0, linetype = "dashed") +
#   # facet_wrap(~fuel_class) + 
#   facet_grid(fuel_class ~ method, scales = "free") +
#   xlim(-5, 10)   +
#   theme(legend.position = "right") +
#   labs(x = "Mean fuel load",
#        y = "Density")   

# The following plot shows the raw (untransformed) values on the left and the transformed values on the right.
# input_class_transform %>%
#   rename(raw_values = value_raw, transformed_values = value) %>%
#   gather(metric, value, raw_values:transformed_values) %>%
#   ggdensity(x = "value", color = "fuel_class") +
#   geom_vline(xintercept = 0, linetype = "dashed") +
#   facet_grid(metric ~ fuel_class, scales = "free") +
#   theme(legend.position = "right", 
#         axis.text = element_text(size = 8),
#         panel.spacing = unit(1, "lines"),
#         legend.text=element_text(size=8),
#         # strip.text = element_text(size = 8),
#         strip.background = element_rect(fill = "gray91", color = "gray91"),
#         panel.border = element_rect(fill = NA, color = "gray91", size = 2)) + 
#   labs(title = "Distribution of values, by fuel class and transformation",
#        caption = "Note: axis scales differ between plots",
#        x = "Mean metric tons per hectare",
#        y = "Density", 
#        color = "Fuel class") + 
#   xlim(-5, 10)
```

##### *Normalize and standardize data*
We applied an arcsine transformation to normalize the plot-level values for total coarse woody debris, calculated as log(x + sqrt(x\^2 + 1)). The arcsine transformation (also called the arcsine square root transformation, or the angular transformation) was identified as the most suitable method using the R function bestNormalize() [bestNormalize package]. Values were standardized upon normalization to have a mean of 0 and standard deviation of 1. (The transformed values for total coarse woody debris were normally distributed at each time point (p \> 0.05), as assessed by Shapiro-Wilk's test. The two outliers (not extreme) detected in the untransformed data were absent after the transformation was applied.)

We subset the plot-level values for coarse woody debris by fuel class, then applied an ordered quantile transformation to normalize the plot-level values for total coarse woody debris, calculated as:

g(x) = psi\^-1 \* ((rank(x) - .5) / (length(x)))

Where psi refers to the standard normal cumulative distribution function, rank(x) refers to each observation's rank, and length(x) refers to the number of observations. The ordered quantile transformation is a rank-based procedure by which the values of a vector are mapped to their percentile, which is then mapped to the same percentile of the normal distribution. Without the presence of ties, this essentially guarantees that the transformation leads to a uniform distribution. Values were standardized by fuel class upon normalization to have a mean of 0 and standard deviation of 1.

The following table lists the transformation applied to each fuel class, as well as the mean and standard deviation (sd) for raw and transformed values. Note the untransformed mean for 1000-hr rotten is \>8, a much larger value compared to those of the other classes that range from 0.162 to 1.016.

------------------------------------------------------------------------

## Total fuel load

### Summary statistics for the plot-level total fuel load by year

The following table summarizes the plot-level total amount of coarse woody debris for each time point.

```{r total_summary}
wd_total %>%
  group_by(lab_time,
           time,
           units) %>%
  get_summary_stats(si_value, type = "mean_sd")  %>%
  mutate(fuel_type = index_wd,
         fuel_class = "All") %>%
  select(fuel_class, time, units, mean, sd, n)   %>%
  fxn_kable()

```

The visual inspection of a box plot of total duff and litter by year showed lower values in post-fire years. Specifically, we observed a decrease in the depth of all duff and litter immediately following the fire (2017; 0-y post-fire). The mean depth of duff and litter did not return to pre-fire levels in subsequent years.

A box plot of total coarse woody debris by year showed lower values in post-fire years. Specifically, we observed a decrease in the depth of all coarse woody debris immediately following the fire (2017; 0-y post-fire). Variance in 2017 was much greater than at any other time point. The mean depth of coarse woody debris did not return to pre-fire levels in subsequent years. However, the fuel load increased from 2017 levels by 2019, and showed little change between 2019 and 2021.

```{r total_boxplot}
wd_total %>%
  ggboxplot(x = "lab_time", 
            y = "si_value", 
            fill = "time", 
            outlier.size = 0.3,
            palette = colors_year) + 
  geom_hline(yintercept = 0, 
             linetype = "longdash", 
             color = "gray50") +
  theme(legend.position = "none", 
        axis.title.x = element_blank()) + 
  scale_x_discrete(labels = function(lab_time) str_wrap(lab_time, width = 20)) + 
  labs(title = paste0(index_wd, " (all)"), 
       y = index_wd_si) 

```

### Main effect of year on total fuel load
We conducted a one-way repeated measures ANOVA to test for a main effect of year on fuel load (i.e., whether total fuel load was significantly different between the four time points).

We found no significant main effect of year on the plot-level total for coarse woody debris (p-adj. = 0.435).

```{r total_main_effect}
wd_total %>%
  mutate(time = lab_time_abbr) %>%
  anova_test(dv = value_norm, 
             wid = plot_id, 
             within = time) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni") %>%
  as_tibble() %>%
  clean_names() %>%
  mutate(method = "me",
         fuel_type = index_wd,
         fuel_class = "All",
         index_value = "value_norm") %>% 
  rename(statistic = f) %>% 
  fxn_signif_adj()  %>%
  select(fuel_class, 
         method, 
         effect, 
         starts_with("p_adj"), 
         statistic,
         starts_with("d"), 
         ges, 
         starts_with("index"))  %>%
  fxn_kable()

```

### Pairwise comparison of total fuel load between years
The result of a post hoc paired comparison between years showed no significant difference in fuel load between years at a significance level of \< 0.05. This finding is consistent with the lack of main effect found for year.

```{r total_pairwise_comparison}
wd_total %>%
  pairwise_t_test(
    value_norm ~ time,
    paired = TRUE,
    p.adjust.method = "bonferroni") %>%
  clean_names() %>%
  mutate(method = "pwc",
         statistic = fxn_digit(statistic),
         fuel_type = index_wd,
         fuel_class = "All",
         index_value = "value_norm") %>%
  fxn_signif_adj() %>%
  select(fuel_class,
         method,
         starts_with("group"),
         starts_with("p_adj"),
         statistic,
         starts_with("d"),
         starts_with("index"),
         - p_adj_signif) %>%
  arrange(p_adj)  %>%
  fxn_kable()

```

------------------------------------------------------------------------

## Mean fuel load, by fuel class

### Summary statistics for the plot-level mean fuel load by fuel class and year
The following table summarizes the plot-level mean values for the five fuel classes at each time point.

```{r class_summary}
wd_class %>%
  # Include time for sorting
  group_by(time, lab_time, units, lab_class, fuel_class) %>%
  get_summary_stats(si_value, type = "mean_sd")  %>%
  mutate(fuel_type = index_wd) %>%
  arrange(fuel_class, time) %>%
  select(lab_class, 
         time, 
         units, 
         mean,
         sd,
         n)  %>%
  fxn_kable()

```

### Visualization of the plot-level mean fuel load by fuel class and year
A box plot of coarse woody debris by year and fuel class showed the post-fire trends differed among the five coarse woody debris fuel classes. [Note: The y-axis scale in the figure below differs by fuel class]

```{r class_boxplot}
input_wd %>%
  # filter(metric %in% "mean") %>%
  ggplot(aes(x = lab_time, 
             y = si_value,  
             fill = lab_time_abbr))  +
  geom_hline(yintercept = 0, 
             linetype = "longdash", 
             color = "gray50") +
  geom_boxplot(outlier.size = 0.8) +
  labs(y = index_wd_si,
       fill = "Timing") + 
  facet_wrap(~fuel_class, scales = "free", nrow = 1) +
  guides(fill = guide_legend(byrow = TRUE)) +
  scale_fill_manual(values = colors_year) +
  theme_fuels_markdown() + 
  theme(legend.position = "top")

# input_class %>%
#   bind_rows(input_total) %>%
#   # Configure factors so total ("all") appears first in facet
#   mutate(fuel_class = ifelse(is.na(fuel_class), "all wd", fuel_class), 
#          lab_fuel = as.character(str_replace(lab_fuel, "hr", "h")), 
#          yr_nn = str_sub(year, 3, 4)) %>%
#   arrange(plot_id, year, fuel_class) %>%
#   mutate(fuel_class = as_factor(fuel_class), 
#          lab_fuel = as_factor(lab_fuel)) %>%
#   ggboxplot(x = "yr_nn", 
#             y = "value", 
#             fill = "lab_year_abbr",
#             outlier.size = 0.3,
#             palette = colors_year) + 
#   theme(legend.position = "top", 
#         axis.text = element_text(size = 8),
#         panel.spacing = unit(0.5, "lines"),
#         # Reduce font size for facet strip text
#         legend.text=element_text(size=8),
#         strip.text = element_text(size = 8),
#         strip.background = element_rect(fill = "gray91", color = "gray91"),
#         panel.border = element_rect(fill = NA, color = "gray91", size = 2)) + 
#   labs(title = paste0(index_wd), 
#        caption = "Note: y-axis scale differs by fuel class", 
#        x = "Year",
#        y = paste0(index_wd_si, " (mean)"), 
#         # Rename legend title
#        fill = "Year") + 
#   facet_wrap(~lab_fuel,  scales = "free", nrow = 1)


```

### Interaction between year and fuel class

We investigated whether there was a significant change in the plot-level fuel load over time when accounting for fuel class. A two-way repeated measures ANOVA was used to determine whether there was a significant interaction between year and fuel class on mean fuel load.

There was a statistically significant two-way interaction between coarse woody debris class and year, F(12, 96) = 5.302, p-adj. = 2.6e-06, ges = 0.242. A significant two-way interaction indicates that the impact of fuel class on fuel load depends on year (and vice versa).

```{r class_interaction}
wd_class  %>%
 anova_test(dv = value_norm, 
             wid = plot_id, 
             within = c(time, fuel_class)) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni") %>%
  as_tibble() %>%
  clean_names() %>%
  mutate(method = "interaction",
         fuel_type = index_wd,
         fuel_class = "By class",
         index_value = "value_norm") %>% 
  rename(statistic = f) %>% 
  fxn_signif_adj()  %>%
  select(method, 
         effect, 
         starts_with("p_adj"), 
         statistic,
         starts_with("d"), 
         ges, 
         starts_with("index")) %>%
  arrange(p_adj) %>%
  filter(effect %in% "time:fuel_class")    %>%
  fxn_kable()


```

#### Effect of treatment on each fuel class

Because we found a significant interaction between fuel class and year, we conducted a main effect assessment for each fuel class.

We found a significant main effect of year on fuel load for 1-hr, 10-hr (p \< 0.001), 100-hr (p \< 0.01), and 1000-hr sound (p \< 0.05) fuel classes. No significant effect of year on fuel load was found for the 1000-hr sound class. One thing the pairwise comparison by fuel class had in common: For the classes with a significant main effect of year on fuel load, the fuel load in 2016 was significantly different from that in 2021.

```{r class_main_effect}
wd_class %>%
  group_by(fuel_class) %>%
  anova_test(dv = value_norm, 
             wid = plot_id, 
             within = time) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni") %>%
  as_tibble() %>%
  clean_names() %>%
  mutate(method = "me",
         fuel_type = index_wd,
         index_value = "value_norm") %>% 
  rename(statistic = f) %>% 
  fxn_signif_adj()  %>%
  select(fuel_class, 
         method, 
         effect, 
         starts_with("p_adj"), 
         statistic,
         starts_with("d"), 
         ges, 
         starts_with("index")) %>%
  arrange(p_adj)   %>%
  fxn_kable()
```

### Pairwise comparisons of mean fuel load between years, by fuel class

We observed a decrease in fuel load immediately after the fire (2017; 0-y post-fire) for all classes of coarse woody debris except 1000-hr sound. Specifically, the fuel load in 2017 for 1-hr, 10-hr, 100-hr, and 1000-hr rotten classes was much lower than pre-fire values; variance was unsurprising. Further, mean values for these four classes remained lower than 2016 values for all post-fire years.

The mean for 1000-hr sound fuels increased immediately after the fire and remained greater than 2016 values for all post-fire years. Variance for this fuel class was elevated in 2017. The temporal trend for all coarse woody debris was more similar to that for the 1000-hr sound fuel class, and showed less similarity to those for other classes. These results suggest that 1000-hr sound fuel was an influential driver of the post-fire patterns observed for all coarse woody debris.

```{r class_pairwise_comparison}
wd_class %>%
  group_by(fuel_class) %>%
  pairwise_t_test(
    value_norm ~ time, 
    paired = TRUE,
    p.adjust.method = "bonferroni") %>%
  clean_names() %>%
  mutate(method = "pwc", 
         statistic = fxn_digit(statistic), 
         fuel_type = index_wd,
         index_value = "value_norm") %>%
  fxn_signif_adj() %>%
  select(fuel_class, 
         method, 
         starts_with("group"),
         starts_with("p_adj"), 
         statistic,
         starts_with("d"), 
         starts_with("index"), 
         - p_adj_signif) %>%
  filter(p_adj < 0.05) %>%
  arrange(fuel_class, p_adj)   %>%
  fxn_kable()
```

**1-hour**

A post hoc pairwise comparison showed a significant difference in 1-hr fuel load between 2016-2017, 2016-2021, and 2017-2019 (p \< 0.01); as well as 2019-2021 (p \< 0.05). No other comparisons were significant.

**10-hour**

A post hoc pairwise comparison showed a significant difference in 10-hr fuel load between 2016-2017 (p \< 0.001), 2016-2021 (p \< 0.01), and 2016-2019 (p \< 0.05). No other comparisons were significant.

**100-hour**

A post hoc pairwise comparison showed no significant difference in 100-hr fuel load between years 2016-2021 at a significance level of \< 0.05.

**1000-hour rotten**

Post hoc comparisons showed no significant difference in 1000-hr rotten fuel load between years at a significance level of \< 0.05.

**1000-hour sound**

Post hoc comparisons showed no significant difference in 1000-hr sound fuel load between years at a significance level of \< 0.05.
